---
title: "Assignment 6 (Module 6)"
author: "Brian Yim"
date: "2025-10-09"
output:
  word_document: default
  html_document: default
---




#Answers: First, when formulating the objective function,
#Each ğ‘¥ğ‘–ğ‘— represents the amount of shipment from the source, plants ğ‘–, to the destination, warehouse ğ‘—, for ğ‘– = 1, 2 and ğ‘— = 1, 2, 3.

#I want to minimize total cost, which includes production cost + shipping cost.
#Production costs: Plant A: $600 per unit Ã— number produced, Plant B: $625 per unit Ã— number produced

#Then, Minimum Z = (600+22)ğ‘¥A1 +(600+14)ğ‘¥A2 + (600+30)ğ‘¥A3 +(625+16)ğ‘¥B1 + (625+20)ğ‘¥B2 + (625+24)ğ‘¥B3

#1.There are constraints from Supply side (i.e., production capacity)
#A1+ ğ‘¥A2+ ğ‘¥A3 â‰¤ 100 (which is the capacity of Plant A)
#B1+ ğ‘¥B2+ ğ‘¥B3 â‰¤ 120 (which is the capacity of Plant B)

#2.There are constraints from Demand side (i.e., monthly demand)
#A1+ ğ‘¥B1 =80 (demand of the warehouse 1)
#A2+ ğ‘¥B2 =60 (demand of the warehouse 2)
#A3+ ğ‘¥B3 =70 (demand of the warehouse 3)

#3.Non-negativity xijâ‰¥0

#next, when we solve this with R

```{r}
# Install lpSolve if not installed
# install.packages("lpSolve")

library(lpSolve)

# Cost matrix (production + shipping)
costs <- matrix(c(
  600+22, 600+14, 600+30,  # Plant A to W1,W2,W3
  625+16, 625+20, 625+24   # Plant B to W1,W2,W3
), nrow=2, byrow=TRUE)

# Supply (production capacity)
supply <- c(100, 120)

# Demand (warehouse)
demand <- c(80, 60, 70)

# Solve transportation problem
result <- lp.transport(
  cost.mat = costs,
  direction = "min",
  row.signs = rep("<=", 2), #plant side constraints
  row.rhs = supply,
  col.signs = rep("=", 3), #warehouse side constraints
  col.rhs = demand
)
```


```{r}
# Show optimal shipping plan
result$solution
```

```{r}

# Show total minimum cost
result$objval
```
#Answer: Minimum Total Cost: $132,790

#Explanation: Plant A produces 100 units and ships 60 to W2 and 40 to W3. Plant B produces 110 units and ships 80 to W1 and 30 to W3. All warehouse demands are met, and total production + shipping cost is minimized.
